---
title: "Individual Assignment 3 Appendix"
output:pdf
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align = "center")
library(ggplot2)
library(tidyverse)
library(dbplyr)
library(ggpubr)
library(plyr)
library(car)
library(ggiraphExtra)
library(zoo)
```
  
  
  
**Part 1:Glimpse of data**
```{r reading dataset, echo=FALSE}
df <- read.csv("complaints-2021-03-14_17_15.csv")
glimpse(df)
```
```{r selecting features, include=FALSE,fig.width = 12, fig.height = 4}

df1 <- df %>% dplyr::select(Date.received,Product) %>%  transmute(Date.received=as.Date(Date.received, format="%m/%d/%y"),Product) %>%  mutate(yearmonth=format(Date.received,"%y%m")) 

df1['yearmonth'] <- type.convert(as.integer(df1$year))
df1<-df1 %>% arrange(Date.received)
```

```{r data filtering,include=FALSE,fig.width = 12, fig.height = 4}
df2<- df1 %>% count('yearmonth')
df3<- subset(df2, yearmonth!="1112")
df4<- subset(df3, yearmonth!="2101")
df5<- subset(df4, yearmonth!="2102")
df6<- subset(df5, yearmonth!="2103")

df7<- dplyr::rename(df6, c("complaints" = "freq"))


# plot1<-ggplot(df13, aes(y=complaints, x=as.factor(yearmonth))) +geom_point() + labs(x="yearmonth",y="complaints")+
# theme(axis.text.x = element_text(size =7, angle = 90))
# plot1


# plot1<- ggplot(aes(x=complaints)) +geom_histogram()
# plot1

plot2 <- ggplot(df7, aes(x=as.factor(yearmonth),y=complaints,fill=as.factor(yearmonth))) + geom_bar(stat = 'identity',show.legend =FALSE)+ labs(x = "Year Month", y = "Complaints", title = "Complaints per Month per Year")+theme(axis.text.x = element_text(angle = 90, hjust = 1))
plot2

```
  
  
   
**Part2:Variance Non-Zero Checking**
```{r asummption checking,echo=FALSE}
# non zero variance checking
vars<-apply(df7, 2, var)
vars
```
   
  
   
**Part 3:Coefficient obtained from linear regression model**
```{r linear regression,echo=FALSE}
model1 <- lm(complaints~yearmonth, data=df7)
model1
```
     
   
   
**Part4:Summary of linear regression model**
```{r summary, echo=FALSE}
summary(model1)
```
  
  
    
**Part 5:Durbin Watson Test Output**
```{r test, echo=FALSE}
durbinWatsonTest(model1)
```
  
  
  
**Part 6: Prediction Calculation**
```{r prediction,echo=FALSE}
values=tibble(yearmonth=c(2201,2201,2203,2204,2205,2206,2207,2208,2209,2210,2211,2212))
predicts<-predict.lm(model1, values)
sum (predicts)
```

  
  
**References**

1. Christiaan Defaux (2019, Aug 9).*Using the Durbin-Watson (DW) test for testing time-series autocorrelation.* Retrieved from
https://medium.com/@cdefaux/using-the-durbin-watson-dw-test-for-time-series-data-b87bd884fbbdhttps://medium.com/@cdefaux/using-the-durbin-watson-dw-test-for-time-series-data-b87bd884fbbd

2. ANDY FIELD, JEREMY MILES, & ZOË FIELD (2012).*Discovering statistics Using r*, SAGE Publications Ltd:My Publisher